---
title: "Prediction"
author: 
  - Huang Lin$^1$
  - $^1$Department of Biostatistics, University of Pittsburgh, 130 De Soto Street, Pittsburgh, PA 15261 USA
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    theme: united
---

```{r setup, include=FALSE}
# rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE, comment = NA,
                      fig.width = 6.25, fig.height = 5)

library(readr)
library(tidyverse) 
library(microbiome)
library(rstatix)
library(corrplot)
library(DT)
options(DT.options = list(
  initComplete = JS("function(settings, json) {",
  "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});","}")))
```

```{r data}
df_merge = read_csv("../data/df_merge.csv")

# OTU table
otu_table = read_tsv("../data/otu-table.tsv")
otu_id = otu_table$`#OTU ID`
otu_table = data.frame(otu_table[, -1], check.names = FALSE, row.names = otu_id)

# Taxonomy table
tax = read_tsv("../data/taxonomy.tsv")
otu_id = tax$`Feature ID`
tax = data.frame(tax[, - c(1, 3)], row.names = otu_id)
tax = tax %>% separate(col = Taxon, 
                       into = c("Kingdom", "Phylum", "Class", "Order", 
                                "Family", "Genus", "Species"),
                       sep = ";")
for (i in 1:ncol(tax)) {
  tax[, i] = sapply(tax[, i], function(x) str_split(x, "__")[[1]][2])
}
tax = as.matrix(tax)
tax[tax == ""] = NA

# Tree
tree = read_tree("../data/tree.nwk")

# Meta data
meta_data = read_tsv("../data/meta_data.tsv")
meta_data$status = factor(meta_data$status, levels = c("nc", "sc"))
meta_data$visit = factor(meta_data$visit, levels = c("v1", "v2"))
rownames(meta_data) = meta_data$sampleid

# Phyloseq object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
sample_names(META) = meta_data$sampleid
TAX = tax_table(tax)
otu_data = phyloseq(OTU, TAX, META, tree)

# Aggregate taxa to the family level
family_data = aggregate_taxa(otu_data, "Family")
family_rarefied = rarefy_even_depth(family_data, rngseed = 1, 
                                    sample.size = 0.9 * min(sample_sums(family_data)), 
                                    replace = FALSE)

# Alpha diversities
alpha_family = alpha(family_rarefied, index = "diversity_shannon")

df_alpha = data.frame(sampleid = meta(family_rarefied)$sampleid,
                      shannon = alpha_family$diversity_shannon)

# Prevotellaceae to Bacteroidaceae
df_pb = abundances(family_data)
df_pb = df_pb[c("Prevotellaceae", "Bacteroidaceae"), ]
df_pb = data.frame(t(df_pb), check.names = FALSE) %>% 
  rownames_to_column("sampleid")

# Family data
df_family = data.frame(t(abundances(family_data)), check.names = FALSE) %>%
  rownames_to_column("sampleid")

df = df_merge %>%
  left_join(df_alpha, by = "sampleid") %>%
  left_join(df_pb, by = "sampleid")

df_wide = df %>%
  pivot_wider(id_cols = c(subjid, status, time2aids, age),
              names_from = visit,
              names_glue = "{visit}_{.value}",
              values_from = il6:Bacteroidaceae)
```

# 1. Log2(%CD4/%CD8) at visit 2 vs. alpha diversity at visit 1

1. Microbial diversity at visit 1 is more negatively correlated with CD4/CD8 among seroconverters than among controls.

2. As the diversity before seroconversion increases, the CD4/CD8 ratio decreases among the serconverters more quickly. This is possibly because controls have a rich diverse ecology of commensal bacteria that are protective, whereas the seroconverters, before seroconversion have a potential loss of commensal bacterial and a possible increase in opportunistic and pathogenic bacteria. As their diversity
Increases, it appears that the CD4/CD8 decreases dramatically for seroconverters. 

3. P-value is obtained from linear regression model adjusting for age.

```{r}
lm_eqn = function(m){
  
  a = unname(coef(m))[1]
  b = unname(coef(m))[2]
  c = unname(coef(m))[3]
  p_val = summary(m)$coef[2, "Pr(>|t|)"]
  
  b = ifelse(sign(b) >= 0, 
             paste0(" + ", format(b, digits = 2)), 
             paste0(" - ", format(-b, digits = 2)))
  c = ifelse(sign(c) >= 0, 
             paste0(" + ", format(c, digits = 2)), 
             paste0(" - ", format(-c, digits = 2)))
  
  eq = substitute(paste(italic(y) == a, b, italic(x), c, italic(age), ", ", italic(p) == p_val),
                  list(a = format(a, digits = 2), b = b, c = c,
                       p_val = format(p_val, digits = 2)))
  
  return(as.character(as.expression(eq)))              
}

df_fig1 = df_wide %>% transmute(subjid, status, age,
                                cd4 = v2_leu3p, cd8 = v2_leu2p,
                                ratio = log(cd4/cd8, base = 2), 
                                shannon = v1_shannon)

df_lm = df_fig1 %>% group_by(status) %>%
  do(fit = lm(formula = shannon ~ ratio + age, data = .)) %>%
  summarise(eq = lm_eqn(m = fit))

df_ann = df_fig1 %>% 
  group_by(status) %>%
  summarise(ratio = 1, shannon = 0) %>%
  mutate(eq = df_lm$eq,
         x = c(0.6, 1), y = c(1.5, 0), status = c("nc", "sc"))

ps8 = ggplot(data = df_fig1, aes(x = shannon, y = ratio, color = status)) + 
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_discrete(name = NULL, label = c("nc" = "NC", "sc" = "SC")) +
  geom_text(data = df_ann, aes(x = x, y = y, label = eq, color = status), 
            parse = TRUE, inherit.aes=FALSE) + 
  labs(x = "Shannon", 
       y = expression(paste(log[2], "(", CD4^{"+"}, "/", CD8^{"+"}, ")")), 
       title = expression(paste(log[2], "(", CD4^{"+"}, "/", CD8^{"+"}, ")",
                                " at Visit 2 vs. Shannon Diversity Index at Visit 1"))) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
ps8
ggsave("../images/supplementary/fig_s8.pdf", height = 5, width = 6.25)
ggsave("../images/supplementary/fig_s8.jpeg", height = 5, width = 6.25, dpi = 300)
```

# 2. Log2(%CD4/%CD8) at visit 2 vs. log2(valerate/propionate) at visit 1

1. Reduction in propionic acid relative to valeric acid reduces CD4 counts relative to CD8 among seroconverters more quickly than among controls.

2. P-value is obtained from linear regression model adjusting for age.

```{r}
df_fig2.1 = df_wide %>% 
  transmute(subjid, status, age,
            cd4 = v2_leu3p, cd8 = v2_leu2p,
            valerate = v1_valerate, propionate = v1_propionate,
            ratio1 = log(cd4/cd8, base = 2), 
            ratio2 = log(valerate/propionate, base = 2))

df_lm = df_fig2.1 %>% 
  group_by(status) %>%
  do(fit = lm(formula = ratio1 ~ ratio2 + age, data = .)) %>%
  summarise(eq = lm_eqn(m = fit))

df_ann = df_fig2.1 %>% 
  group_by(status) %>%
  summarise(ratio = 1, shannon = 0) %>%
  mutate(eq = df_lm$eq,
         x = c(-4, -4), y = c(1.5, -0.5), status = c("nc", "sc"))

p = ggplot(data = df_fig2.1, aes(x = ratio2, y = ratio1, color = status)) + 
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_discrete(name = NULL, label = c("nc" = "NC", "sc" = "SC")) +
  geom_text(data = df_ann, aes(x = x, y = y, label = eq, color = status), 
            parse = TRUE, inherit.aes = FALSE) + 
  labs(x = expression(paste(log[2], "(Valerate/Propionate)")), 
       y = expression(paste(log[2], "(", CD4^{"+"}, "/", CD8^{"+"}, ")")), 
       title = expression(paste(log[2], "(", CD4^{"+"}, "/", CD8^{"+"}, ") at Visit 2 vs. ",
                                log[2], "(Valerate/Propionate) at Visit 1"))) +
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
p
```

# 3. Microbiome

## 3.1 Log2(prevotellaceae/bacteroidaceae) at visit 1 vs. status.

1. Absolute abundance of prevotellaceae relative to bacteroidaceae at visit 1 is highly correlated with seroconversion at a future visit.

2. P-value is obtained by ANCOVA test adjusting for age.

```{r}
df_fig3.1_raw = df_wide %>% 
  transmute(subjid, status, age, 
            prev = v1_Prevotellaceae,
            bact = v1_Bacteroidaceae,
            ratio = log(prev/bact, base = 2)) %>%
  filter(is.finite(ratio))

df_fig3.1 = df_fig3.1_raw %>% 
  group_by(status) %>%
  summarise(ratio = mean(ratio, na.rm = TRUE))

tab3.1 = df_fig3.1
col_name = c("Status", "Log2(Prevotellaceae/Bacteroidaceae)")
colnames(tab3.1) = col_name
tab3.1 %>% 
  datatable(caption = "Log2(Prevotellaceae/Bacteroidaceae) at Visit 1 vs. Status") %>%
  formatRound(col_name[-1], digits = 2)

df_jitter = df_fig3.1_raw %>%
  filter(ratio >= quantile(ratio, 0.1, na.rm = TRUE) & 
           ratio <= quantile(ratio, 0.9, na.rm = TRUE))

df_ann = data.frame(x = 1.3, 
                    y = max(df_jitter$ratio, na.rm = TRUE),
                    y_max = 1.3 * max(df_jitter$ratio, na.rm = TRUE),
                    status = "nc", ratio = 1, 
                    p = anova_test(data = df_fig3.1_raw, 
                                   formula = ratio ~ status + age)$p[1]) %>%
  mutate(label = paste0("p = ", signif(p, 2)))

p = ggplot(data = df_fig3.1, aes(x = status, y = ratio)) + 
  geom_col(color = "black", aes(fill = status)) + 
  geom_jitter(position = position_jitter(0.2), data = df_jitter, 
              alpha = 0.5, size = 0.8) +
  labs(x = NULL, y = NULL, 
       title = expression(paste(log[2], "(Prevotellaceae/Bacteroidaceae) at Visit 1 vs. Status"))) +
  scale_fill_discrete(name = NULL, 
                      label = c("nc" = "NC", "sc" = "SC")) +
  geom_blank(data = df_ann, aes(y = y_max)) +
  geom_label(data = df_ann, aes(x = x, y = y, label = label),
             size = 4, vjust = -0.5, hjust = 0, color = "black") +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
p
```

## 3.2 Correlations between significant families and pro-inflammatory cytokines at visit 1

```{r}
df_fig3.2_raw = df %>% 
  dplyr::select(-Prevotellaceae, -Bacteroidaceae) %>%
  left_join(df_family, by = "sampleid") %>%
  filter(visit == "v1") %>%
  transmute(status, CD14 = cd14, CD163 = cd163, `IL-6` = il6, LBP = lbp,
            Succinivibrionaceae, `S24-7`, `[Mogibacteriaceae]`,
            Coriobacteriaceae, Erysipelotrichaceae, `[Odoribacteraceae]`,
            Verrucomicrobiaceae, Bacteroidaceae, `[Barnesiellaceae]`, 
            Rikenellaceae)

df_fig3.2_nc = df_fig3.2_raw %>%
  filter(status == "nc") %>%
  dplyr::select(-status) %>%
  mutate_if(is.numeric, scale)

df_fig3.2_sc = df_fig3.2_raw %>%
  filter(status == "sc") %>%
  dplyr::select(-status) %>%
  mutate_if(is.numeric, scale) %>%
  mutate_if(is.numeric, scale)

# NC
mat_cor_nc = round(cor(df_fig3.2_nc, use = "complete.obs"), 2)

corrplot(mat_cor_nc, method = "circle") %>%
  corrRect(index = c(1, 5, 14))

# SC
mat_cor_sc = round(cor(df_fig3.2_sc, use = "complete.obs"), 2)

corrplot(mat_cor_sc, method = "circle") %>%
  corrRect(index = c(1, 5, 14))
```

# 4. Prediction model

1. Age, alpha diversity, log ratio of prevotellaceae to bacteroidaceae, and log ratio of valerate to propionate at visit 1 predicts seroconversion at a future visit.

```{r}
df_fig4 = df_wide %>% 
  transmute(subjid, status, age, 
            shannon = v1_shannon,
            vale = ifelse(v1_valerate == 0, NA, log(v1_valerate, base = 2)), 
            prop = ifelse(v1_propionate == 0, NA, log(v1_propionate, base = 2)),
            prev = ifelse(v1_Prevotellaceae == 0, NA, log(v1_Prevotellaceae, base = 2)),
            bact = ifelse(v1_Bacteroidaceae == 0, NA, log(v1_Bacteroidaceae, base = 2)),
            ratio1 = prev - bact,
            ratio2 = vale - prop) %>%
  mutate(prob = ifelse(status == "sc", 1, 0))
df_fig4$status = factor(df_fig4$status)
```

## 4.1 Model 1: status (NC vs. SC) ~ age + alpha diversity + log2(prevotellaceae/bacteroidaceae) + log2(valerate/propionate)

```{r}
fit1 = glm(status ~ age + ratio1 + ratio2 + shannon, 
           family = "binomial", data = df_fig4) 
tab4.1 = rstatix::tidy(fit1)
tab4.1$term = c("Intercept", "Age", "Log2(Prevotellaceae/Bacteroidaceae)",
                "Log2(Valerate/Propionate)", "Shannon")
tab4.1 %>% datatable(caption = "Model 1") %>%
  formatRound(colnames(tab4.1)[-1], digits = 3)
```

## 4.2 Model 2: status (NC vs. SC) ~ age + alpha diversity + log2(prevotellaceae/bacteroidaceae)

```{r}
fit2 = glm(status ~ age + ratio1 + shannon, family = "binomial", data = df_fig4) 
tab4.2 = rstatix::tidy(fit2)
tab4.2$term = c("Intercept", "Age", "Log2(Prevotellaceae/Bacteroidaceae)", "Shannon")
tab4.2 %>% datatable(caption = "Model 2") %>%
  formatRound(colnames(tab4.2)[-1], digits = 3)
```

## 4.3 Model 3: status (NC vs. SC) ~ age + log2(prevotellaceae/bacteroidaceae)

```{r, fig.height=5, fig.width=10}
fit3 = glm(status ~ age + ratio1, family = "binomial", data = df_fig4) 
tab4.3 = rstatix::tidy(fit3)
tab4.3$term = c("Intercept", "Age", "Log2(Prevotellaceae/Bacteroidaceae)")
tab4.3 %>% datatable(caption = "Model 3") %>%
  formatRound(colnames(tab4.3)[-1], digits = 3)

df_fig4_long = df_fig4 %>% 
  dplyr::select(subjid, prob, age, ratio1) %>%
  gather(key = "measure", value = "value", age:ratio1)
df_fig4_long$measure = recode(df_fig4_long$measure,
                              age = "Age", 
                              ratio1 = "Log2(Prevotellaceae/Bacteroidaceae)")
df_fig4_long2 = df_fig4_long %>%
  filter(measure == "Log2(Prevotellaceae/Bacteroidaceae)")

df_ann = data.frame(measure = "Log2(Prevotellaceae/Bacteroidaceae)", 
                    x = 1, y = 0.7, prob = 1, 
                    p = coef(summary(fit3))["ratio1", "Pr(>|z|)"]) %>%
  mutate(label = paste0("p = ", signif(p, 2)))

p = df_fig4_long2 %>% ggplot(aes(x = value, y = prob)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(x = expression(paste(log[2], "(Prevotellaceae/Bacteroidaceae)")), 
       y = "Probability of Seroconversion", 
       title = "Probability of Seroconversion Predicted by Logistic Regression Model") +
  geom_label(data = df_ann, aes(x = x, y = y, label = label),
             size = 4, vjust = -0.5, hjust = 0, color = "black") +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.background = element_rect(fill = "white"),
        plot.title = element_text(hjust = 0.5))
p
```

# Session information

```{r, message = FALSE, warning = FALSE, comment = NA}
sessionInfo()
```


